\documentclass[a4paper,10pt]{letter}
\usepackage[utf8x]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{Doug Imbruce}
\address{Joshua Bowles}

\opening{Re: Taxonomist Consultant}


My interest in establishing a long-term consulting relationship with Qwiki is twofold: I believe in start-up companies, and I believe in the application of taxonomies and ontologies to knowledge representation in the real world. I am a young, motivated scholar interested in producing commercially viable applications from theoretical investigations. My BA and MA degrees have been done in traditional linguistics programs; while my PhD will be done in the newer field of computational linguistics. To this end, I am looking to establish strong industry ties while I pursue my doctorate; and in the longer run, hope to bring my academic research into the real world.

I can bring to Qwiki my spirit of innovation balanced with my sense of reliance on methods that work. My experience in working with taxonomies from various fields (such as semantic hierarchies for language parsers, language family classification, noun-classifiers in various natural languages, and ontologies in artificial intelligence and formal pragmatics) gives me a broad base to work from. I can also bring to Qwiki the discipline and drive needed to carry out large scale research and detail oriented results. 

I look forward to hearing from you. 


\signature{Joshua Bowles}

\closing{\textsl{Thank you for your time and consideration},}

%enclosure listing
%\encl{}

\end{letter}
\end{document}
