% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=9in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Joshua W. Bowles}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left
 
 
\begin{letter}{Lingua \\
2563 152nd Ave NE\\
Suite C\\
Redmond, Washington 98052\\
{}\quad\\
c.c. Native English (US), Linguistics, and Software}


\begin{flushleft}
{\large\bf Joshua W. Bowles}
\end{flushleft}
\medskip\hrule height 1pt
\begin{flushright}
\hfill (801) 420-0884\\ 
\hfill \texttt{bowlesling@gmail.com}\\ 
\end{flushright} 
\vfill % forces letterhead to top of page
 

\opening{Dear Job Search Committee,}
\noindent I am very interested in applying to the native English (US) speaker position at your company. 

\noindent I have an expertise in linguistics with a special focus in syntax, semantics, and pragmatics. My graduate school specialization was in the morphosyntax of native American Indian languages and comparative linguistic theories. After graduating I obtained an internship position as an industry linguist at Attensity Corporation where I was able to put into practice much of the theory I had been taught. I was also able to begin acquiring real-world computational skills relevant to the consumer market.   

\noindent My formal education focused on documenting endangered languages and designing text grammars for language learners and scholars. To these ends I have expertise in evaluating and analyzing a wide-range of linguistic phenomena from a diverse set of languages; this skill has also made me a better analyst of English. I also engaged in high level research into formal and mathematical logic, specializing in recursive procedures. Since graduation I have been focusing on computational linguistics. This includes using a variety of programming languages for specific goal-oriented projects. The languages I have been most involved with are Python (and the Natural Language Toolkit), Scheme, Java, and C++. Software projects include chatbots for language instruction in minority languages, a spell-checker for a minority language, chatbots to deal with idioms and context-dependent information, and a number of language corpora for specific linguistic phenomena in varieties of English.  

\noindent I offer Lingua my expertise in linguistic analysis, my experience in putting theory to practice, and my determination to get it right. I have a deep committment to computational linguistics and an unswerving drive to learn all that I can.  

\noindent I can bring to Lingua a committment to excellence and a spirit of collaboration. I look forward to hearing from you.

  

\closing{Sincerely,} 
 
\end{letter}
 

\end{document}
