\documentclass[a4paper,10pt]{letter}
\usepackage[utf8x]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{Twitter}
\address{Joshua Bowles}

\opening{Re: Open Source Advocate}

Twitter is cool. Scala is awesome. Ruby is great. Clojure (and Storm) are freakin' rad! Since starting to learn Clojure and Scala I've been reading through alot of the Twitter open source projects and all I can say is cool! Twitter is cool. Open source is still the future.

I came to software development from academia. While shopping for PhD programs in computational linguistics I took a summer position at a company doing research and ended up staying on as software engineer. I'm now on the core team responsible for the company's primary transactional system. During my time I've been able to particiapte in many architectural design projects and have taken part in helping to open source the result of some of these projects. I also experiment frequently and always open source the code. Lately, I've been working on personal open source projects that I intend to release as production-ready.

I am a developer, first and foremost. I enjoy building, learning, and doing research, for which I spend a great deal of personal time. I also enjoy sharing the products of developers' passion. I am still starting out in my career and as I become a better developer over time I anticipate building and maintaining more open source projects. I feel that at this point in my career a role as an open source advocate would be greatly beneficial. I am excited by what Twitter has done and is doing. I am personally experimenting with and building tools on top of many of the open source projects Twitter has released and am familiarizing myself with those tools more everyday; it would be awesome to do this as part of the Twitter team.



\closing{\textsl{Thank you for your time and consideration},  \signature{Joshua Bowles}}

%enclosure listing
%\encl{}

\end{letter}
\end{document}
