\documentclass[11pt]{article}

\usepackage[usenames]{xcolor}


\begin{document}
\author{Joshua Bowles}
\title{Eng 2020 Fall 2009 Schedule}
\date{\today}


\maketitle

   {\bf The schedule here is {\it tentative} and will most likely be modified. It is your responsibility to keep track of the changes as I announce them in class.}\\
   \begin{enumerate}
\item AUG 26 W: (week 1): Introductions; find syllabus, register UV link and e-mail; \textbf{Read Scholar I, DK:204}
\item[] AUG 28 F: Open discussion about convention, audience, why we write; \textbf{Read A\&B ch. 9}

\item  AUG 31 M (week 2): Library: critical observation (Scholar I) 
\item[] SEP 2 W: \textcolor{red}{\texttt{Library assignment (Scholar I) DUE;}} critical reading A\&B:115-17; EXERCISE---A\&B:139\ldots thesis statement  
\item[] SEP 4 F: Research questions A\&B:243-5; A\&B:579, 602-4; \textbf{Read Scholar II}

\item  SEP 7 M (week 3): \textcolor{magenta}{LABOR DAY HOLIDAY}
\item[] SEP 9 W: Thesis statement\ldots bring on Friday; \textbf{Read DK:102-3,146-7.}
\item[] SEP 11 F: Meet in library; \textcolor{red}{\texttt{Response to Scholar II DUE;}} EXERCISE---A\&B:223\ldots work on extending thesis statement; \textbf{Read DK:104-113}

\item  SEP 14 M (week 4): Begin paper \#1
\item[] SEP 16 W: What is evidence? EXERCISE---A\&B:44\ldots evidence to support thesis statements; \textbf{Read A\&B:41-47}
\item[] SEP 18 F: Essay \#1 peer review; \textbf{Read DK:88-96}

\item SEP 21 M (week 5): Writing day
\item[] SEP 23 W:  \textcolor{red}{\texttt{\#1 Informative and Surprising (4-6 pgs) DUE;}} What is Analysis/Synthesis; find two articles for Essay \#2 
\item[] SEP 25 F:  EXERCISE---A\&B:350 1-6; {\bf Read A\&B ch. 13, DK:2-13}

\item SEP 28 M (week 6): What is Rhetoric?
\item[] SEP 30 W: Essay \#2 peer review (1st)
\item[] OCT 2 F: EXERCISE---A\&B:359\ldots rhetorical strategies; finish A\&B ch. 13

\item OCT 5 M (week 7): Research day
\item[] OCT 7 W: \textcolor{red}{\texttt{Summaries to research articles (2 pgs; 1 per article) DUE;}} EXERCISE---A\&B:360\ldots similarities and differences
\item[] OCT 9 F: Essay \#2 peer review (2nd)

\item OCT 12 M (Week 8): Research and citation
\item[] OCT 14 W:  Visit Library: interlibrary loan; database search; \textcolor{red}{\texttt{\#2 Analysis and Synthesis (4-6 pgs) DUE;}}  EXERCISE---A\&B:362\ldots generating ideas for rhetorical strategies {\bf Read DK:18-87}--(as much as
you can over the break)
\item[]  OCT 15 F: \textcolor{magenta}{FALL BREAK (begins on Thursday)} 

\item OCT 19 M (week 9): Library day DO YOUR RESEARCH!! (also look at part 8 of DK esp. page 360)
\item[] OCT 21 W: \textcolor{red}{\texttt{Response to Y DUE;}} drafting proposals and abstracts; bring a
proposal next time!
\item[] OCT 23 F: Peer review of abstracts (\#3)

\item OCT 26 M (week 10): \textcolor{red}{\texttt{\#3 Abstract (1 pg) and Annotated Bibliography DUE}}; {\bf Read A\&B ch. 14}
\item[] OCT 28 W: Essay\#4 peer review (1st)
\item[] OCT 30 F: Essay \#4 peer review (1.1); start making presentations (use draft as guideline)

\item NOV 2 M (week 11): {\bf Read A\&B ch. 15} 
\item[] NOV 4 W: The logic of argument---Parsons
\item[] NOV 6 F: More on logic

\item NOV 9 M (week 12): Light week; {\bf Re-read A\&B chs. 14, 15}
\item[] NOV 11 W: \#4 peer review (2nd)
\item[] NOV 13 F: A\&B ch.15

\item NOV 16 M (week 13): Research Presentations (5-10 minutes)
\item[] NOV 18 W: Research Presentations (5-10 minutes)
\item[] NOV 20 F: Research Presentations (5-10 minutes)

\item NOV 23 M (week 14):  
\item[] NOV 25 W: \textcolor{magenta}{THANKSGIVING HOLIDAY}
\item[] NOV 27 F: \textcolor{magenta}{THANKSGIVING HOLIDAY}

\item NOV 30 M (week 15): Essay \#4 peer review (3rd); \textcolor{red}{\texttt{Response Z DUE}}; directions for \textcolor{blue}{Portfolio}
\item[] DEC 2 W: EXERCISE In-class workshop---supporting claims with evidence
\item[] DEC 4 F: EXERCISE In-class workshop---style, voice, audience

\item DEC 7 M (week 16): Essay \#4 peer review (4th and last one)
\item[] DEC 9 W:  Open discussion about logic, style, voice, audience---will accept Portfolios
\item[] DEC 11 F: STUDY DAY

\item DEC 14 M (FINALS WEEK!!): open
\item[] DEC 16 W: \textcolor{blue}{\texttt{Portfolio, Classical Argument (10-12pgs) DUE}}
\item[] DEC 17 F: \textcolor{green}{FALL 2009 SEMESTER ENDS}
\end{enumerate}

\end{document}