\documentclass[11pt]{article}

\usepackage[usenames]{xcolor}


\begin{document}
\author{Joshua Bowles}
\title{Eng 1010 Fall 2009 Schedule}
\date{\today}


\maketitle

{\bf The schedule here is {\it tentative} and will most likely be modified. It is your responsibility to keep track of the changes as I announce them in class.}
   \begin{enumerate}
\item AUG 26 W: (week 1): Introductions; find syllabus, register UV link and e-mail
\item[] AUG 28 F: Open discussion about convention, audience, why we write; \textbf{Read A\&B chs. 1, 2;} library assignment.

\item  AUG 31 M (week 2): Library: critical observation
\item[] SEP 2 W: \textcolor{red}{\texttt{Library assignment DUE;}} What is a: Question?, Claim? Problem? EXERCISE---You claim, You assume 
\item[] SEP 4 F: What is a Problem?; begin paper \#1; {\bf Read DK 26-27} 

\item  SEP 7 M (week 3): \textcolor{magenta}{LABOR DAY HOLIDAY}
\item[] SEP 9 W: Problem, Context, Presupposition? EXCERCISE---You assert, You assume; {\bf Read DK 28-33} 
\item[] SEP 11 F: Meet in library: database search---find a scholarly article for Paper \#3

\item  SEP 14 M (week 4): Problem/Topic Paper
\item[] SEP 16 W: \textcolor{red}{\texttt{\#1 Problematizing a Topic (2 pgs) DUE;}} peer review grading; {\bf Read A\&B: chs. 4, 16}
\item[] SEP 18 F: What is a Solution? Research strategies; begin paper \#2 

\item SEP 21 M (Week 5): Research day in library
\item[] SEP 23 W: Essay \#2 peer review (1st)
\item[] SEP 25 F: Essay \#2 peer review (1.1)

\item SEP 28 M (week 6): Localizing a problem 
\item[] SEP 30 W: \textcolor{red}{\texttt{\#2 Solving a Local Problem (4-6 pgs) DUE;}} {\bf Read A\&B: chs. 6, 21}
\item[] OCT 2 F: EXERCISE---summarize and respond 

\item OCT 5 M (week 7): Writing day 
\item[] OCT 7 W: \textcolor{red}{\texttt{\#3 Summary and Strong Response (4-6 pgs) DUE;}} EXERCISE---drafting proposals or abstracts; bring a proposal next time!
\item[] OCT 9 F: Peer review proposals \#4; {\bf Read A\&B: chs. 8, 23}
 
\item OCT 12 M (Week 8): Research, A\&B ch. 23; DK part 8
\item[] OCT 14 W: Visit Library: interlibrary loan; database search; {\bf Read DK:18-87} (as much as you can!)
\item[]  OCT 15 F: \textcolor{magenta}{FALL BREAK begins on Thursday} 

\item OCT 19 M (week 9): Library day DO YOUR RESEARCH!! (look at part 8 of DK esp. page 360)
\item[] OCT 21 W:  A\&B ch. 8
\item[] OCT 23 F: \textcolor{red}{\texttt{\#4 Annotated Bibliography \& Proposal DUE;}} 

\item OCT 26 M (week 10): More on exploration and citation
\item[] OCT 28 W: A\&B ch. 22
\item[] OCT 30 F: A\&B ch. 23

\item NOV 2 M (week 11): {\bf look at A\&B ch. 15} 
\item[] NOV 4 W: The logic of exploring---argument and information
\item[] NOV 6 F: More on logic

\item NOV 9 M (week 12): Light week; {\bf Re-read A\&B chs. 22, 23, and possibly 15}
\item[] NOV 11 W: \#5 peer review (1st)
\item[] NOV 13 F: \#5 peer review (1.1); start making presentations (use draft as guideline)

\item NOV 16 M (week 13): Research Presentations (5-10 minutes)
\item[] NOV 18 W: Research Presentations (5-10 minutes)
\item[] NOV 20 F: Research Presentations (5-10 minutes)

\item NOV 23 M (week 14):  
\item[] NOV 25 W: \textcolor{magenta}{THANKSGIVING HOLIDAY}
\item[] NOV 27 F: \textcolor{magenta}{THANKSGIVING HOLIDAY}

\item NOV 30 M (week 15): Essay \#5 peer review (2nd); \textcolor{red}{\texttt{Response Z DUE}}; directions for \textcolor{blue}{Portfolio}
\item[] DEC 2 W: EXERCISE In-class workshop---supporting claims with evidence
\item[] DEC 4 F: EXERCISE In-class workshop---style, voice, audience

\item DEC 7 M (week 16): Essay \#5 peer review (3rd and last one)
\item[] DEC 9 W:  Open discussion about logic, style, voice, audience---will accept Portfolios
\item[] DEC 11 F: STUDY DAY

\item DEC 14 M (FINALS WEEK!!): open
\item[] DEC 16 W: \textcolor{blue}{\texttt{Portfolio, Exploratory Research (10-12pgs) DUE}}
\item[] DEC 17 F: \textcolor{green}{FALL 2009 SEMESTER ENDS}
\end{enumerate}

\end{document}