\begin{cv}{CV --- Joshua Bowles}
\hfill
   
\begin{cvlist}{Contact Information}
	\item[Company] \flexref{http://www.go1on1.com/}{One On One Marketing}\\
				2912 Executive Pkwy,\\
				Suite 300,\\ 
				Lehi, UT 84043
\item[Mail]     \begin{verbatim}bowlesling@gmail.com, jbowles@1on1.com\end{verbatim}
\item[Github]			\flexref{http://jbowles.github.com}{jbowles-Github} 
\item[Bitbucket]              \flexref{https://bitbucket.org/jbowles}{jbowles-Bitbucket}
\item[Technical Blog]      \flexref{http://jbowles.github.com/}{jbowles-lambda} 
\end{cvlist}

\begin{cvlist}{Overall Experience}
\item[Since 2007] Research in mathematical, biological, documentary, and computational linguistics.
\item[Since 2010] Research and software development in business.
\item[Since 2013] Research in various machine learning techniques.
\item[Current/Future] Familiarizing myself with tools using large data sets and machine learning techniques for text processing; deliver near real-time results for large data analysis (for example, Go and Ruby web apps with complex and accurate text classification for near real-time twitter stream); experiment with different distributed system architectures; experiment with different concurrency models (via Scala, Ruby, Clojure, Go) and the kinds of data problems suited to them.
\end{cvlist}

\begin{cvlist}{Employment}
\item[2010 -- present] {\bf Software Developer and Researcher}, One On One Marketing\\
			Internet Marketing Company, Lehi, UT,\\
			\textsl{Develop and maintain in-house products as well as lead-gen transaction and validation system with Ruby and Rails;  Maintain production transaction system in Ruby and co-ordinate with Business Intelligence; Technical lead on re-architecting, writing, and delivering production release of new distribution system uisng Ruby and Nodejs; Co-architect of new transaction system in Ruby and in the process of porting it to Go on personal time.}
	
\item[2008 -- 2010] {\bf Adjunct Lecturer}, Department of English\\
     Utah Valley University, Orem, UT.\\
		 \textsl{Research techniques, informal logic, and pragmatics relevant to science composition.}

\item[2009, July -- Nov] {\bf Computational Linguist (Intern)}, Attensity Corporation\\
     U.S. Technology Center, Salt Lake City, UT.\\
		\textsl{Knowledge engineering; named entity recognition; compiling and judging corpora of lexical
		items designed to fit within specific ontologies; some work on developing methods for introducing
		fixed phrases and idioms into the lexicon.}  
\end{cvlist}

\begin{cvlist}{Consulting Work}
\item[2013, 9 months] Utah, USA. Contextual classification of (non)standard unicode text in social media.
\item[2007, 4 months] Northern Rio Negro, Brazil. Endangered Language Preservation Techniques with FOIRN (Federac\~ao das Organizac\~oes Indigenas do Rio Negro). 
\item[2004, 4 months] Colorado, USA. Afghani Poshtun sources for language instruction in English and Poshtun.
\end{cvlist}

\begin{cvlist}{Education}
% \item \begin{cvlist}{}
                \item[Present] PhD on hold. \\ Personal work in Computational Linguistics, Machine Learning, and Computer Science  
                \item[2006--2008] \href{http://www.linguistics.utah.edu/}{University of Utah}\\
                        M.A., Theoretical Linguistics\\
                        Thesis:
												\href{http://content.lib.utah.edu/cdm/singleitem/collection/etd2/id/1431/rec/1}{\emph{Agreement in Tuyuca}}\\
												Advisor: Lyle Campbell 
                \item[2000--2006] \href{http://www.msudenver.edu/las/}{Metropolitan State University of Denver}\\
                        B.A., English\\ 
			(minor in Linguistics; completed 24 credit hours in Philosophy)\\
                        Honours Thesis: \emph{Dislocated Poetics: Language, Biology, Brain, and Literature}
												Advisor: Paul Farkas 
%       \end{cvlist}
\end{cvlist}

\begin{cvlist}{Open Source - Personal}
\item[Go - current focus]{\href{http://jbowles.github.io/rivet/}{Rivet}: Natural Language Processing tools}
\item[Ruby - on hold]{\href{http://jbowles.github.com/ruth/}{Ruth}: Stream, encrypt, and parse application logs using Distributed Ruby. This will be pulled out to 3 separate projects in the future.}
\item[Ruby -  on hold]{\href{https://github.com/jbowles/taml_hub}{Taml Hub}: Experimenting with twitter stream, text classification, sentiment analysis, and Hadoop.}
\item[Clojure - on hold]{\href{https://github.com/jbowles/entropy}{Entropy}: Using MaxEnt models from Java's OpenNLP to experiment with parsing unstructured text from application log files}.
\end{cvlist}

\begin{cvlist}{Open Source - Contributed}
\item[Ruby]{\href{https://github.com/hopsoft/ellington}{Ellington}: Simple state-machine inspired architecture for streamlining transactional processing. {\bf Significant role in designing and making architectural decisions; product of redesigning transaction system at current job.}}
\item[Ruby]{\href{https://github.com/hopsoft/grand}{Grand}: Extract-Transform-Load MySQL tables to csv. {\bf Helped with requirements, architecting, and production implementation.}}
\item[Ruby]{\href{http://hopsoft.github.com/goldmine/}{Goldmine}: MapReduce type functionality using Ruby. {\bf Early adopter and tester.}}
\end{cvlist}


\begin{cvlist}{Programming Languages}
\item[Most Experience] Ruby (1.8x, 1.9.x, 2.x), Go (1.x)
\item[Favorites] Ruby, Go, Lisp
\item[Actively Learning] Go, Nodejs, Clojure, Scala, Church\\ \tiny{(in order of emphasis)}
\end{cvlist}

\begin{cvlist}{Research}
\item[General] {\bf Computational and Theoretical Linguistics, Software Development}
		\small
\item[\small Linguistic] Evidentiality, Idioms \& Polysemy, Modality, Tense, Context 
\item[\small Other] Logic, Probability, Algebra, Artificial Intelligence, Machine Learning
		\normalsize
\item[Projects on Hold] \textcolor{Mygrey}{Polysemy and Context: What is Computational Semantics Good For?\\ Recursion in Language\\ Symmetry in X-bar Structure: Graphs and Algebraic Groups in Linguistic Theory\\ Syntax and Pragmatics of Evidentiality in English\\ Corpus of English Evidentiality Data (CEED)}
%\texttt{Indefinitely suspended so I can foucs on math and programming:}\\
%\textcolor{Mygrey}{}
\end{cvlist}

\begin{cvlist}{Storage \& Transport Technologies}
\item[Most Experience] MySQL, PostgreSQL
\item[Actively Learning] Kafka, Doozer, Storm, Hadoop (and ecosystem)
\item[Limited] MongoDB, Redis, Oracle, Hadoop
\item[Interests] Riak, Neo4j
\end{cvlist}

\begin{cvlist}{Other Random Technologies}
\item[Markup] \LaTeX, HTML, MarkDown
\item[Languages] SQL, PLSQL, Prolog 
\item[Tools] Emacs, Vim, Oracle Application Express, JIRA, Virtual Box
\item[OS] Linux (Ubuntu, Fedora, Arch, CentOS), Mac OSX, Windows 
\end{cvlist}

\begin{cvlist}{Natural Languages}
\item English (native),  Spanish (basic speaking and reading), Portuguese (basic reading, beginning
speaking), German and Old English (1 year formal study each), Tuyuca and Desano (structural and research knowledge), Afghani Poshto (academic knowledge), General knowledge of many different natural languages and their properties. 
\end{cvlist}

\begin{cvlist}{Field Work}
\item[2007, 4 months] Sao Gabriel da Cachoeria and Camanaus, Brazil. Work on all grammatical
aspects of Desano and limited work on Cubeo; Tukanoan family. 
\item[2004, 1 month] Denver, Colorado, USA. Limited work on phonology of Afghani Poshto;
Indo-Iranian family.
\end{cvlist}

\begin{cvlist}{Academic Grants}
\item[2007] Endangered Language Fund (Co-PI) --- \$3,500  
\item[2007] International Field Work Travel Funds, University of Utah --- \$300
\end{cvlist}

\begin{cvlist}{Publications}
\item[2009] The morphosyntax of tense-evidentials: An initial model.
\emph{\href{http://vanhise.lss.wisc.edu/ling/?q=node/21}{LS0} Working Papers in Linguistics: Proceedings
of the 6th Workshop in General Linguistics,} Volume 8. University of Wisconsin-Madison.

\item[2009] The Eastern Tukanoan languages and the typology of classifiers. \emph{Proceedings of the 33rd
Annual Meeting of the \href{http://linguistics.berkeley.edu/BLS/past_meetings.html}{Berkeley
Linguistics Society.}} (Wilson Silva and Joshua Bowles)

%----------------
\item[2009] Review of Heck (2008), \emph{On pied-piping: Wh-movement and beyond} (Mouton de
Gruyter). \href{http://linguistlist.org/issues/20/20-2283.html}{LinguistList}.

%---------------
\item[2008] Some questions about determining causal inference and criteria for evidence:
Response to Ladd, Dediu, and Kinsella (2008).
\emph{\href{http://www.biolinguistics.eu.}{Biolinguistics}} 2.2-3: 247-255.

%---------------
\item[2007] Review of Niyogi (2006), \emph{The computational nature of language learning and
evolution} (MIT Press). {\it Word} 57.2: 178-183. \texttt{(pre-published version
\href{http://sites.google.com/site/bowleslinguistics/Home/research/teaching-1/publications}{here})}

%--------------
\item[submitted] Notice of Booij (2007), \emph{The grammar of words} (Oxford University Press).
\href{http://www.elanguage.net/home.php}{eLanguage.}

\item Notice of Sch\"afer (2008), \emph{The syntax of (anti-)causatives: External arguments in
change-of-state contexts} (Amsterdam: John Benjamins).
\href{http://www.elanguage.net/home.php}{eLanguage.}

\item Language families. \emph{Encyclopedia of Native American History} ed. by
\href{http://college.usc.edu/faculty/faculty1003494.html}{Peter C.Mancall}. New York: Facts on File.
\texttt{(see submitted draft
\href{http://sites.google.com/site/bowleslinguistics/Home/research/teaching-1/publications}{here})}

\item Linguistic areas. \emph{Encyclopedia of Native American History} ed. by Peter C. Mancall. New
York: Facts on File. \texttt{(see submitted draft
\href{http://sites.google.com/site/bowleslinguistics/Home/research/teaching-1/publications}{here})}
\end{cvlist}

\begin{cvlist}{Talks \& Conferences}
\item[2008] Evidentials, Agreement, and Tuyuca: Initial Hypotheses. \emph{Linguistics Department
Student Conference} 7, University of Utah, April 10.

\item Evidentiality and Agreement Interactions in Tuyuca: A Generative Approach.
\emph{Workshop in General Linguistics} 6, University of Wisconsin-Madison, April 4-5.

\item Agreement and Tense-Evidentiality Morphs in Tuyuca. \emph{Workshop on Structure
and Constituency in the Languages of the Americas} 13, Queens University,
Ontario, Canada, March 28-30 (\texttt{unable to attend}).

\item The ``Language of Observation'' in Linguistics (Informal Presentation). University of Utah
Linguistics Reading Group, February 15.
%-----------------
\item[2007] Neural correlates of Grammatical Gender: Functionalist and Formalist
Assumptions (Informal Presentation). University of Utah Linguistics Reading
Group, October 15.

\item Classifiers and Typology in Eastern Tukanoan Languages. \emph{7th Biennial Meeting of the
Association for Linguistic Typology}, Paris, France, September 25-28.
(Wilson Silva and Joshua Bowles)

\item Formalism and Functionalism in Language Documentation. \emph{Department of
Foreign Languages and Literatures}, Universidade Federal do Amazonas, Brazil,
June 10.

\item The Amazonian Languages Research and Documentation Group. Poster. \emph{3rd
Annual Conference on Endangered Languages and Cultures of the Native
Americas}, University of Utah, April 13-15.
(Wilson Silva and Joshua Bowles)

\item The Eastern Tukanoan Languages and the Typology of Classifiers. \emph{33rd Annual Berkeley
Linguistics Society Conference}, University of Berkeley, California,
February 9-11. (Wilson Silva and Joshua Bowles)

\item A Typology of Nominal Classifiers in the Eastern Tukanoan Languages.
\emph{Department of Linguistics Speakers Series}, University of Utah, January 24. (Wilson Silva and
Joshua Bowles)
%--------------------
% \item[2006] Dislocated Poetics: Towards a Neurobiological Explanation for Poetry and Poetic
% Language. \emph{Honors Faculty Colloquium}, Metropolitan State College of Denver,
% Colorado, May 05.
%--------------------
\item[2004] Neurobiology, Poetry, and Dislocation. \emph{English Faculty Colloquium}, Metropolitan
State College of Denver, Colorado, April 25. (Paul Farkas and Joshua Bowles)
%--------------------
\item[2002] Philosophy of Language, the Dialectic Method, and James Joyce. \emph{Honors Regional
Conference}, University of Colorado, Boulder, March 24-25.
\end{cvlist}

\begin{cvlist}{Teaching}

\item[2010/2008]        \emph{Science \& Technology Writing and Research} ENG 2020, Undergraduate\\ 
                           Utah Valley University (9 courses total)
\item[Summer 2009]      \emph{Humanities \& Social Science Writing and Research} ENG 2010, Undergraduate\\
                           Utah Valley University\\
                        \emph{Introduction to Composition} ENG 1010, Undergraduate\\ 
                           Utah Valley University
\item[Spring 2008/2007] Teaching Assistant, \emph{Introduction to Language} LING 1200, Undergraduate\\
                        University of Utah, David Iannucci (1 semester), Mauricio Mixco (2 semesters)
\item[Fall 2006]        Teaching Assistant, \emph{Cross-cultural Communication} LING 3600, Undergraduate\\
                           University of Utah, Stephen Sternfeld                        
\end{cvlist}


\begin{cvlist}{Academic Services}
\item[2007-2008] Volunteer work for \textsl{Conference on Endangered Languages and Cultures of the
North Americas}, University of Utah.
\item[2007] Attended by invitation with Wilson Silva the 20th Anniversary Meeting of FOIRN
(Federac\~ao das Organizac\~oes Indigenas do Rio Negro) Taracua, Brazil: June 25-29.  
\item[2006] Co-founding member of research group: Amazonian Languages Research and Documentation,
Center for American Indian Languages, University of Utah. 
\item[2003] Lead student organizer and introductory comments for ``Noam Chomsky: Live at the Auraria
Campus,'' Presented by MSCD Student Activities, Denver, Colorado, April 04.
\end{cvlist}

\end{cv}
