The past is a strange beast. I had a dream last night about two people I once knew, Angus Hicks and Jerry Payne, and when I woke up I wanted to see them. I know both of these guys from my time in Denver. My Denver time is only time in life I felt like I truly made friends. I miss those guys.

For some time I have wanted to take you girls to Denver. I want to take you to eat at a restuarant. It was at this restuarant that I made my friends. Jerry Payne, Angus HIcks, Don Hudson, Teri Rippetto. They all revolve around ``the restaurant".

When the Fall season comes around I get remembrances from the past. Smells, the light of day, the feel of air... they all bring the past back to me. The past is dangerous becuase it bids us to reside in it as long as we can. But there is no moving forward in the past, and the remembrance of the past is never accurate. It is colored by the feeling of today. Sometimes the past is emberrasing, sometimes comforting. If remebering the past is everything, that is, if it includes a full spectrum of emotion the you are doing it right. If the past is nothing more than the same wave repeating itself then something is wrong.

The past is a paradox. It never changes. It always changes. This is what it means to be a human. If the past is stored in bytecode then it never changes. The solid state memory of an event never changes. The signature of an event written into magnetic disk never changes. But humans are not bytecode or solid state drives or magentic disks. We are not 800 millimeter tape, we are not flash drives. Events captured without interpretation are static waves repeating. But this doesn't dampen the paradox. When I look at pictures of you girls when you were 9 years or 2 years or 6 months I feel again, and again differently each time colored by the passage of time. And when I remember you... that too is colored by the passage time... by the transformation of flesh and bone and synapse.

I feel this for my own personal past. I miss many things and wish that many things were not there. But I remember it all anyway, I make myself remember the emberrasements and the mistakes just as much as the highlights and joys. I remember working at Potager and the friends I made. It was the only time in my adult like that I felt like I had real friends. Uncharacteristically of me, I miss them. I do not usually miss people, I do not usually want to see poeple after my time with them is done. You girls, your mother, and friends from Denver: I miss.