%HANDOUT

\documentclass[11pt]{article}


\usepackage{natbib,graphicx,times,linguex,cgloss4e,latexsym}
\usepackage{qtree}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

\begin{document}
\author{Joshua Bowles}
\title{Statement of Purpose: Brandeis Computational Linguistics}
\date{\today}


\maketitle



\documentclass[11pt]{article}
\usepackage{natbib,graphicx,times,linguex,cgloss4e,latexsym}
\usepackage{qtree}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

\begin{document}
\author{Joshua Bowles}
\title{Statement of Purpose: Stanford}
\date{\today}


\maketitle
\section{Personal}
I am looking to change research focus from documentary lingusitics to more formal and computational techniques. This means that I probably do not have the particular educational bakcground of most people going into computational lingusitics, but my committment and passion are as strong as anybody's.

I was successful in previous MA studies, having produced or co-produced work that was accepted to high level conferences (see CV). However, after two years of typical lingistics training---including specializations in the morphology and syntax of native American languages in both a typological and Minimalist (generative) setting---I realized that I had become increasingly interested in more rigorous approaches. Capitalizing on my undergraduate training in mathematical logic I started studying modern algebra. I acquired an internship as a linguist working on computational linguistics and liked the work (see CV). Presently I am studying core mathematics and programming laguages needed for computational linguistics tasks and begun building my own corpora and experimenting with it. I have even applied for a National Endowment of Humanities Summer Stipend to work full time on the project. 

At this point, I need to someone to take a chance on me and provide me the training I need. I do not have appropriate background in computer science or mathematics to apply to typical natural language processing or AI programs. And honestly, I don't want to give up my lingusitics training but would, instead, rather apply it directly to the language problems in NLP. (Additionally, there has been a growing sentiment (e.g., ) in the NLP field that more language-centric study is needed and I do have a very strong linguistics background which I would like to apply to the NLP sciences.) 

\subsection{GRE Scores}
I realize that my GRE scores are not good and that this does not reflect well on my merit as a student. If you measure the merit of an active mind by a static remainder of standardized expectations then I will not measure up. Putting aside the rhetoric, I assure you that the GRE is not an accurate reflection of my potential and that my true worth is metted out in the daily grind of evaluating techinques, ideas, and data: these cannot be tested. I can also assure you that although I, apparently, do not test well, my committment to acquiring the mathematical and computational background necessary for being successful in the field of natural language investigation is solid.    


\end{document}  