\documentclass[line]{resume}

\usepackage{graphicx,wrapfig}
\usepackage[names]{xcolor}
\definecolor{bolinkcolor}{rgb}{.4,.2,.0}
\usepackage[pdftex,colorlinks,breaklinks,
		linkcolor=bolinkcolor,
		citecolor=bolinkcolor,
		urlcolor=bolinkcolor,
		plainpages=false,
		pdfpagelabels,
		bookmarks=false,
		pdfstartview=FitH]{hyperref}
\setlength{\topmargin}{-0.6in}  % Start text higher on the page 
\setlength{\textheight}{9.8in}  % increase textheight to fit more on a page
\setlength{\headsep}{0.2in}     % space between header and text
\setlength{\headheight}{12pt}   % make room for header
\usepackage{fancyhdr}  % use fancyhdr package to get 2-line header
%\renewcommand{\headrulewidth}{0pt} % suppress line drawn by default by fancyhdr
\lhead{\ttfamily \scriptsize Joshua Bowles} % force lhead all the way left
\rhead{\ttfamily \scriptsize Page \thepage}  % put page number at right
\chead{\ttfamily \scriptsize Last Updated \today}
\cfoot{}  % the footer is empty
\pagestyle{fancy} % set pagestyle for the document
\thispagestyle{empty} % this page does not have a header

\begin{document} 
{\sc \Large {\bf Short CV --- Joshua W. Bowles}}
\begin{resume}


\section{Selected Publications}

(submitted) Language families. \emph{Encyclopedia of Native American History} ed. by \href{http://college.usc.edu/faculty/faculty1003494.html}{Peter C.Mancall}. New York: Facts on File. \texttt{(see submitted draft \href{http://sites.google.com/site/bowleslinguistics/Home/research/teaching-1/publications}{here})}

(submitted) Linguistic areas. \emph{Encyclopedia of Native American History} ed. by Peter C. Mancall. New York: Facts on File. \texttt{(see submitted draft \href{http://sites.google.com/site/bowleslinguistics/Home/research/teaching-1/publications}{here})}

(in press) The Eastern Tukanoan languages and the typology of classifiers. \emph{Proceedings of the 33rd
Annual Meeting of the \href{http://linguistics.berkeley.edu/BLS/past_meetings.html}{Berkeley Linguistics Society.}} (Wilson Silva and Joshua Bowles)

(in press) The morphosyntax of tense-evidentials: An initial model. \emph{\href{http://ling.wisc.edu/lso/wpl-main.html}{LS0} Working Papers in Linguistics: Proceedings of the 6th Workshop in General Linguistics,} Volume 8. University of Wisconsin-Madison.

(2009) Review of Heck (2008), \emph{On pied-piping: Wh-movement and beyond} (Mouton de
Gruyter). \href{http://linguistlist.org/issues/20/20-2283.html}{LinguistList}.

(2008) Some questions about determining causal inference and criteria for evidence:
Response to Ladd, Dediu, and Kinsella (2008). \emph{\href{http://www.biolinguistics.eu.}{Biolinguistics}} 2.2-3: 247-255.

(2007) Review of Niyogi (2006), \emph{The computational nature of language learning and evolution} (MIT Press). {\it Word} 57.2: 178-183. \texttt{(pre-published version \href{http://sites.google.com/site/bowleslinguistics/Home/research/teaching-1/publications}{here})}

\section{Selected Talks and Conferences}
(2008a) Evidentiality and Agreement Interactions in Tuyuca: A Generative Approach.
\emph{Workshop in General Linguistics} 6, University of Wisconsin-Madison, April 4-5.

(2008b) Agreement and Tense-Evidentiality Morphs in Tuyuca. \emph{Workshop on Structure
and Constituency in the Languages of the Americas} 13, Queens University,
Ontario, Canada, March 28-30 (\texttt{unable to attend}).

(2007a) Classifiers and Typology in Eastern Tukanoan Languages. \emph{7th Biennial Meeting of the Association for Linguistic Typology}, Paris, France, September 25-28.
(Wilson Silva and Joshua Bowles)

(2007b) Formalism and Functionalism in Language Documentation. \emph{Department of
Foreign Languages and Literatures}, Universidade Federal do Amazonas, Brazil,
June 10.

(2007c) The Amazonian Languages Research and Documentation Group. Poster. \emph{3rd
Annual Conference on Endangered Languages and Cultures of the Native
Americas}, University of Utah, April 13-15.
(Wilson Silva and Joshua Bowles)

(2007d) The Eastern Tukanoan Languages and the Typology of Classifiers. \emph{33rd Annual Berkeley Linguistics Society Conference}, University of Berkeley, California,
February 9-11. (Wilson Silva and Joshua Bowles)

(2004) Neurobiology, Poetry, and Dislocation. \emph{English Faculty Colloquium}, Metropolitan State College of Denver, Colorado, April 25. (Paul Farkas and Joshua Bowles)

\section{Programming}
Actively learning: Python, Scheme, and Prolog; intersts in Haskell, C\# 


\section{Field Work}
{\bf 2007, June-August} Sao Gabriel da Cachoeria and Camanaus, Brazil. Work on all grammatical aspects of Desano and limited work on Cubeo; Tukanoan family. {\bf 2004, September} Denver, Colorado, USA. Limited work on phonology of Afghani Poshtun; Indo-Iranian family.


 \end{resume}
\end{document}





























